Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

### 2025-01-22 manuscript/appendix.qmd
Compare sample to CPS

---
title: 'Appendix for "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"'
short-title: 'Appendix for "Why Donors Donate"'
date: "2025-01-20"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
crossref: 
  tbl-title: "Table&nbsp;A"
  fig-title: "Figure&nbsp;A"
toc: true
toc-depth: 2
format: 
  hikmah-pdf:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
  hikmah-manuscript-pdf+ms:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
---

@@

tar_load(c(
  participant_summary, model_summary_table, feature_lookup, level_lookup
))

@@

```{r}
#| label: tbl-sample-details-demographics
#| tbl-cap: Respondent demographics
#| tbl-pos: H
#| include: true
tbl_participant_summary <- participant_summary %>% 
  select(
    category, 
    Question = clean_name,
    Response = level,
    N = count,
    `%` = nice_prop
  ) %>% 
  split(~category)

@@

>>>>

---
title: 'Appendix for "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"'
short-title: 'Appendix for "Why Donors Donate"'
date: "2025-01-20"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
crossref: 
  tbl-title: "Table&nbsp;A"
  fig-title: "Figure&nbsp;A"
toc: true
toc-depth: 3
format: 
  html: default
  hikmah-pdf:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
  hikmah-manuscript-pdf+ms:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
---

@@

tar_load(c(
  participant_summary, model_summary_table, feature_lookup, level_lookup, cps_diffs
))

@@

```{r}
#| label: tbl-sample-details-demographics
#| tbl-cap: Respondent demographics
#| tbl-pos: H
#| include: true
tbl_participant_summary <- participant_summary$participant_summary %>% 
  mutate(level = case_match(
    level,
    "Age: Less than median" ~ "Less than 2017 national median (36)",
    "Income: Less than median" ~ "Less than 2017 national median ($61,372)",
    .default = level
  )) %>%
  mutate(level = str_remove(level, "^[^:]+: ")) %>%
  select(
    category, 
    Question = clean_name,
    Response = level,
    N = count,
    `%` = nice_prop
  ) %>% 
  split(~category)

@@

```{r}
#| label: tbl-sample-cps
#| tbl-cap: Sample characteristics compared to nationally representative Current Population Survey (CPS) estimates
#| include: true

rows_with_dagger <- cps_diffs %>%
  mutate(row = row_number()) %>%
  filter(cps_prop_outside_ci == TRUE) %>%
  pull(row)

notes <- list(
  `*` = list(i = 0, j = 4, text = tt_md("Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.")),
  `†` = list(i = rows_with_dagger, j = 3, text = tt_md("National value is substantially different from the sample; the 95% posterior credible interval for the difference between the sample and national proportions contains 0.")),
  a = list(i = 1:5, j = 1, text = "Annual Social and Economic Supplement (ASEC) of the Current Population Survey (CPS), March 2019"),
  b = list(i = 6:7, j = 1, text = "Monthly CPS, September 2019"),
  c = list(i = 8, j = 1, text = "Monthly CPS, November 2018")
)

cps_diffs %>%
  mutate(across(c(prop_sample, national_value,), fmt_pct)) %>%
  mutate(diff_ci = glue::glue(
    "{fmt_pp_int(y)}\n[{fmt_pp_int(ymin)}, {fmt_pp_int(ymax)}]"
  )) %>%
  select(
    Variable = variable,
    Sample = prop_sample,
    National = national_value,
    `∆` = diff_ci
  ) %>%
  mutate(`∆` = tt_linebreak(`∆`)) %>%
  tt(width = c(0.45, 0.1, 0.1, 0.15), notes = notes) %>%
  format_tt(j = 4, escape = FALSE) %>%
  format_tt(j = 1:3, escape = TRUE) %>%
  style_tt(j = 1, align = "l") %>%
  style_tt(j = 2:4, align = "c") %>%
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = "rowsep=2pt, cells={font=\\footnotesize}, row{1}={valign=b}"
  )
```

### 2025-01-22 manuscript/manuscript.qmd
Compare sample to CPS

We fielded our survey experiment through Centiment, which recruits representative samples of paid (and highly engaged) survey participants online. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question about their philanthropic behavior early in the survey---if a participant responded that they give once every few years or never, they were disqualified from the study. Our reported causal effects are therefore not representative of the entire population and instead are only valid for people interested in charitable giving. We also included an attention check question early in the survey and removed respondents who failed the question. Importantly, these screening questions were presented *prior to* the experimental manipulation to avoid post-treatment bias [@MontgomeryNyhanTorres:2018a]. Following screening, we received 1,016 viable responses. Tables A1, A2, and A3 provide a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across all pre-treatment characteristics, including gender, age, education, income, and attitudes toward charity.

@@

The issue area an organization works on also serves as a heuristic for donors. As seen in the marginal means in @fig-all-results, organizations focused on human rights and emergency response are more popular than those working on issues related to the environment or refugee relief. When using emergency response (the most popular issue) as the reference category for AMCEs, working with human rights causes a `r amce$human_rights$pp_abs` percentage point decrease (`r amce$human_rights$nice_ci_pd`), while environmental and refugee relief issues see `r amce$environment$pp_abs` (`r amce$environment$nice_ci_pd`) and `r amce$refugee_relief$pp_abs` (`r amce$refugee_relief$nice_ci_pd`) percentage point decreases, respectively. To put these effects in context, these effects are smaller than the the Red Cross effect and a little larger than the Oxfam and Greenpeace effects.

>>>>

We fielded our survey experiment through Centiment, which recruits representative samples of paid (and highly engaged) survey participants online. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question about their philanthropic behavior early in the survey---if a participant responded that they give once every few years or never, they were disqualified from the study. We also included an attention check question early in the survey and removed respondents who failed the question. Importantly, these screening questions were presented *prior to* the experimental manipulation to avoid post-treatment bias [@MontgomeryNyhanTorres:2018a]. Following screening, we received 1,016 viable responses. Tables A1, A2, and A3 provide a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across all pre-treatment characteristics, including gender, age, education, income, and attitudes toward charity. Additionally, Table A4 compares sample characteristics with nationally representative estimates from the 2019 Current Population Survey (CPS). Our sample matches national proportions of age, marital status, and education, and contains slightly more men and slightly higher income levels than the general population. Due to the initial screening question that limited participants to those with an interest in charitable activity, respondents are far more likely to have donated, volunteered, or voted in the past year than population CPS estimates. Our reported causal effects are therefore not representative of the entire population and instead are only valid for people interested in charitable giving.

@@

The issue area an organization works on also serves as a heuristic for donors. As seen in the marginal means in @fig-all-results, organizations focused on human rights and emergency response are more popular than those working on issues related to the environment or refugee relief. When using emergency response (the most popular issue) as the reference category for AMCEs, working with human rights causes a `r amce$human_rights$pp_abs` percentage point decrease (`r amce$human_rights$nice_ci_pd`), while environmental and refugee relief issues see `r amce$environment$pp_abs` (`r amce$environment$nice_ci_pd`) and `r amce$refugee_relief$pp_abs` (`r amce$refugee_relief$nice_ci_pd`) percentage point decreases, respectively. To put these effects in context, these effects are smaller than the Red Cross effect and a little larger than the Oxfam and Greenpeace effects.
